---
title: "WUP_04"
author: "Rita Jarbanda!"
date: "fall 2023"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
mydata <- read.csv(file="GSScsv.csv")
```

```{r eval=FALSE}
mydata <- mydata %>%
mutate(across(where(is.character), as.factor))
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

Using ChatGPT and Candp helper files to compare numerical data and categorical data in the GSS data set. We will compare WORDSUM and how it effected by Education level.

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
ggplot(mydata, aes(x = DEGREE, y = WORDSUM))+
  geom_boxplot()+
  labs(
    x = "DEGREE",
    y = "WORDSUM",
    title = "Scatter plot of WORDSUM by DEGREE"
  )+
  theme_minimal()
```

The plot indicates that there is a notable difference in wordsum scores between those who are high school educated and all those groups who are college educated. There is no notable difference in wordsum scores between those who are junior college educated and those with their bachelors degree. There is also a substantial gap between those who have their graduate degrees, with the graduate degree groups 50th percentile is above the median for every other group.


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r include=TRUE, echo=FALSE}
favstats(WORDSUM~DEGREE,data=mydata)
```

There is a general trend in the means of the different groups that moves upwards as the level of education increases. Those who are junior college educated fall outside of this trend with a mean wordsum score that lies far above what could be expected from the general trend in means.


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=mydata)
summary.aov(model1)
```

The very small p-value of 0.00236 indicates there is a statistically significant difference in wordsum scores depending on educational level.

```{r}
plot(TukeyHSD(model1, conf.level = 0.98),las=2, cex.axis = 0.3)
```

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
